---
title: 'Reproducible Research: Peer Assessment 1'
output:
  pdf_document: default
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data

Load activity data.
```{r}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activity <- read.csv('activity.csv')
```

Cobines intervals into date time column.
```{r}
time <- formatC(activity$interval / 100, 2, format='f')
activity$date.time <- as.POSIXct(paste(activity$date, time),
                                 format='%Y-%m-%d %H.%M',
                                 tz='GMT')
```

Converts the dates to be for today.
```{r}
activity$time <- format(activity$date.time, format='%H:%M:%S')
activity$time <- as.POSIXct(activity$time, format='%H:%M:%S')
```


## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values

Calculate the mean for each day.
```{r}
total.steps <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```

Mean and mediam for each day:
```{r}
mean(total.steps)
median(total.steps)
```

Histogram of distribution of total number of steps per day:
```{r histogram}
library(ggplot2)
qplot(total.steps, xlab='Total steps', ylab='Frequency',type="1")
```


## Are there differences in activity patterns between weekdays and weekends?
