---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data

Load activity data.
```{r}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activity <- read.csv('activity.csv')
```

Cobines intervals into date time column.
```{r}
time <- formatC(activity$interval / 100, 2, format='f')
activity$date.time <- as.POSIXct(paste(activity$date, time),
                                 format='%Y-%m-%d %H.%M',
                                 tz='GMT')
```

Converts the dates to be for today.
```{r}
activity$time <- format(activity$date.time, format='%H:%M:%S')
activity$time <- as.POSIXct(activity$time, format='%H:%M:%S')
```


## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values

Calculate the mean for each day.
```{r}
total.steps <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```

Mean and mediam for each day:
```{r}
mean(total.steps)
median(total.steps)
```

Histogram of distribution of total number of steps per day:
```{r histogram}
library(ggplot2)
qplot(total.steps, xlab='Total steps', ylab='Frequency',type="1")
```



## What is the average daily activity pattern?

Calculate the mean steps for each five minute interval, and then put it in a data frame.
```{r}
mean.steps <- tapply(activity$steps, activity$time, mean, na.rm=TRUE)
daily.pattern <- data.frame(time=as.POSIXct(names(mean.steps)),
                            mean.steps=mean.steps)
```

Plot the mean steps.
```{r timeseriesplot}
library(scales)
ggplot(daily.pattern, aes(time, mean.steps)) + 
    geom_line() +
    xlab('Time of day') +
    ylab('Mean number of steps') +
    scale_x_datetime(labels=date_format(format='%H:%M'))
```

Which five minute interval has the highest mean number of steps?
```{r}
most <- which.max(daily.pattern$mean.steps)
format(daily.pattern[most,'time'], format='%H:%M')
```


## Imputing missing values
Identify the number of intervals with missing step counts ("NA's"):
```{r}
summary(activity$steps)
```

Fill missing values, using means.
```{r}
library(Hmisc)
activity.imputed <- activity
activity.imputed$steps <- with(activity.imputed, impute(steps, mean))
```

Compares the mean and median steps for each day between the original data set and the imputed data set.
```{r}
total.steps.imputed <- tapply(activity.imputed$steps, 
                              activity.imputed$date, sum)
mean(total.steps)
mean(total.steps.imputed)
median(total.steps)
median(total.steps.imputed)
```

And a histogram of the imputed dataset.
```{r histogram_imputed}
qplot(total.steps.imputed, xlab='Total steps', ylab='Frequency')
```

Imputing the missing data has increased the average number of steps. 

## Are there differences in activity patterns between weekdays and weekends?

Add a factor column for whether a day is a weekday or weekend.
```{r}
day.type <- function(date) {
    if (weekdays(date) %in% c('Saturday', 'Sunday')) {
        return('weekend')
    } else {
        return('weekday')
    }
}

day.types <- sapply(activity.imputed$date.time, day.type)
activity.imputed$day.type <- as.factor(day.types)
```

Dataframe holding the mean steps for weekdays and weekends.
```{r}
mean.steps <- tapply(activity.imputed$steps, 
                     interaction(activity.imputed$time,
                                 activity.imputed$day.type),
                     mean, na.rm=TRUE)
day.type.pattern <- data.frame(time=as.POSIXct(names(mean.steps)),
                               mean.steps=mean.steps,
                               day.type=as.factor(c(rep('weekday', 288),
                                                   rep('weekend', 288))))
```

Compares the patterns between weekdays and weekends.
```{r timeseries_daytype}
ggplot(day.type.pattern, aes(time, mean.steps)) + 
    geom_line() +
    xlab('Time of day') +
    ylab('Mean number of steps') +
    scale_x_datetime(labels=date_format(format='%H:%M')) +
    facet_grid(. ~ day.type)
```
